---
title: "Introduction to Data Visualisation"
subtitle: "NZLS Workshop Day"
author:
  name: "Elena Sheard"
  affiliation: "Te Kāhui Roro Reo | New Zealand Institute of Language, Brain and Behaviour <br/> Te Whare Wānanga o Waitaha | University of Canterbury"
format:
  revealjs:
    theme: [custom.scss]
    incremental: true
    logo: images/NZILBB-small.svg
    template-partials:
      - title-slide.html
    title-slide-attributes:
      # If you have Marsden funding, change image name to `nzilbb-uc-marsden.svg`
      data-background-image: images/nzilbb-uc.svg
      # First number controls the horizontal position, second controls vertical.
      data-background-position: '50% 5%'
      # Controls size of image relative to width of the slide.
      data-background-size: 70%
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="icon" type="image/png" sizes="32x32" href=".//images/fav.png" />
editor: 
  markdown: 
    wrap: 72
execute: 
  echo: true
knitr:
  opts_chunk: 
    dev: "png"
    dev.args: 
      bg: "transparent"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
qb_vowels <-read_rds(here('Data','qb_vowels.rds')) 
```

## Workshop Overview

-   Introduction to data visualisation with ggplot2 package and its main
    functions
-   How to make three plots
    -   Density, scatter and box plots
-   How to customise these plots
    -   Colour, labels , legends, themes, faceting
-   Additional tips (saving plots, useful packages)

## How it will work

-   Go to https://github.com/ems1010/NZLS-ggplot2-workshop
    -     Fork (if you know how) OR
    -     Download as .zip folder `<> Code` button -\> `Download ZIP`
-  Unzip folder, open the file **NZLS-ggplot2-workshop.Rproj** in RStudio
    -   File -\> Open Project -\> Where you have saved the folder
-   Contains everything we will need

## The project

-   In the R project I have provided:
    -   a .rds file of data from the QuakeBox (from the `nzilbb.vowels`
        package)
    -   A Quarto document with activities
    -   The code for this presentation (RMarkdown file)
-   Open the Quarto script **NZLS_Workshop_ggplot2_Activities** in the
    *Files* tab (bottom right corner)

## Some reassurance

-   If I can learn how to make graphs in R, you can too!

## What is ggplot2?

-   ggplot2 is a R package dedicated to data visualisation\
-   Has an underlying grammar ('grammar of graphics')
    -   Lets you build graphs by combining independent components
    -   The same underlying data can be transformed by many different
        scales or layers
    -   This allows you to build almost any type of chart (flexible!)
-   It is also over a decade old, meaning there are a lot of resources
    available

## Using ggplot2

```{r,echo=TRUE, eval=FALSE}
install.packages("RColorBrewer") 
library(RColorBrewer)
library(tidyverse)
qb_vowels <-read_rds(here('Data','qb_vowels.rds')) 
```

-   To use ggplot2, you need to:
    -   Install and load either the `ggplot2` package or the `tidyverse`
        package
    -   Load the data
-   I have also included code for installing and loading the
    RColorBrewer package

## The Basics: data and mapping

-   All plots are composed of:
    -   The **data**: the information you want to visualise
    -   A **mapping**: a description of how you want the variables in
        your data to be 'mapped' to aesthetic attributes like colour,
        shapes, or x and y axes

## The Basics continued

-   All plots are also composed of:
    -   The **layers**: the visualisation(s) that you make/see,
        displayed by **geoms** (geometrical objects)
    -   The **coordinates** on the plot axes
    -   **Scales** which specify the aesthetic mapping (make the points
        *this* colour)

## Building a plot

-   All plots you make in ggplot2 will begin with the `ggplot()`
    function
    -   This builds the first component of your graph (the base layer)\
-   You also need to tell ggplot what data you want to visualise
    -   The name of your dataframe or object, in our case `qb_vowels`

## An overview of the data

-   Monologic data from the QuakeBox
    -   Male and female speakers
    -   Range of ages and ethnic backgrounds
-   NZE monophthong F1/F2 measurements

## Categorical variables (finite categories)

-   age (participant_age_category)
-   ethnicity (participant_nz_ethnic)
-   word
-   gender (participant_gender)
-   following_segment (following_segment_category)
-   speaker
-   vowel

## Continuous variables

-   duration (vowel_duration)
-   F1 midpoint (F1_50)
-   F2 midpoint (F2_50)
-   articulation rate (articulation_rate)
-   word frequency (word_freq)
-   amplitude

## The Base Layer

```{r,eval=TRUE}
ggplot(data=qb_vowels)
```

## ggplot2: Mapping

-   Depends on what kind of graph you are after, but for *most* you will
    want to add **x and/or y axes**
-   Within `ggplot(data=qb_vowels)`, we need another function `aes()`
    within which you give the x and y axes so that:
    -   ggplot(data=*dataframe*, aes(x=*column_a*, y=*column_b*)) 
    -   The names you give for 'x' and 'y' are the **names of the
        columns in the dataframe you want to plot**

## ggplot2: Mapping

```{r, eval=TRUE}
ggplot(data=qb_vowels, aes(x=F2_50,y=F1_50))
```

## ggplot2: building a scatter plot

-   We already have the basis for a scatterplot as we have **two
    continuous variables** on the x and y axes
-   To turn this into a scatterplot, add `+ geom_point()` to the
    previous code
    -   This tells ggplot2 that we want to build a scatter plot, with
        the specified x and y axes
-   When you add a new component to a graph, there must be a `+`
    connecting to the previous one
-   And voila!

## Basic Scatter plot

```{r,eval=TRUE}
ggplot(data=qb_vowels, aes(x=F2_50,y=F1_50))+
  geom_point()
```

## Mini-Task 1: Build a scatter plot

-   Make your own scatter plot by changing the x and y axes to different
    continuous variables from the dataframe:
    -   duration (vowel_duration)
    -   F1 midpoint (F1_50)
    -   F2 midpoint (F2_50)
    -   articulation rate (articulation_rate)
    -   word frequency (word_freq)
    -   amplitude

## Building a box plot

-   The basic `ggplot(data=, aes(x=,y=))` stays the same
    -   Instead of `+geom_point()` we add `+geom_boxplot()`
-   Box plots use:
    -   a **categorical** variable on the **x axis**
    -   a **continuous** variable on the **y axis**

## Basic box plot

```{r}
ggplot(data=qb_vowels, aes(x=vowel,y=F1_50))+ 
  geom_boxplot()
```

## Mini-Task 2: Build a box plot

-   Make your own box plot by changing the y axis to a different
    continuous variables from the dataframe:
-   And change the x axis to a different categorical variable from the
    dataframe

## Building a density plot

-   The basic code is `ggplot(data=, aes(x=))`
    -   This time we add `+geom_density()`
-   Density plots use:
    -   a **single continuous** variable on the **x-axis**
    -   Y axis tells you the distribution of this variables

## Basic density plot

```{r}
ggplot(data=qb_vowels, aes(x=F1_50))+
  geom_density()
```

## Mini-Task 3: Build a density plot

-   Make your own density plot by changing the x axis to a different
    continuous variable from the dataframe
    -   duration (vowel_duration)
    -   F1 midpoint (F1_50)
    -   F2 midpoint (F2_50)
    -   articulation rate (articulation_rate)
    -   word frequency (word_freq)
    -   amplitude

## Customising in ggplot2: Colour

-   Colour is a very easy way to add additional information
-   We map colour to variables within the `aes()` function, after we
    have put in the x and y axes
    -   `ggplot(data=data_frame, aes(x=columm_a,y=column_b,color=column_c))`
-   The same principle applies to shape and linetype

## Scatter Plot with Colour

```{r,}
ggplot(data=qb_vowels, aes(x=F2_50,y=F1_50, color=vowel))+
  geom_point()
```

## Customising in ggplot2: Colour

-   If your graph has shapes with lines (like boxplots and density
    plots), you can:
    -   change the line colour with `color=column_c`
    -   change the colour inside the lines with `fill=column_d`

## Boxplot with colour

```{r, eval=FALSE}
qb_vowels %>%
  filter(vowel %in% c("DRESS", "KIT")) %>%
  ggplot(aes(
    x = vowel,
    y = F1_50,
    fill = participant_nz_ethnic,
    color = participant_gender
  )) +
  geom_boxplot()
```

## Boxplot with colour

```{r, echo=FALSE}
qb_vowels %>%
  filter(vowel %in% c("DRESS", "KIT")) %>%
  ggplot(aes(
    x = vowel,
    y = F1_50,
    fill = participant_nz_ethnic,
    color = participant_gender
  )) +
  geom_boxplot()
```

## Mini-task 4: Customise with colour

-   Customise a scatter plot `geom_point()`, box plot (`geom_boxplot()`)
    and density `geom_density()` plot with colour
-   If you are after a challenge: filter/summarise data going into the
    plot first

## Labels and Titles

-   The `+labs()` function lets you change the plot title, caption, x
    and y axes and the legend labels
    -   `labs(title="Plot Title",caption="Plot Caption",x="column_a",y="column_b", color="column_c",fill="column_d")`

## Changing our figure title

```{r,eval=FALSE, echo=TRUE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = vowel)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )
```

## Labels and Titles

```{r,eval=TRUE, echo=FALSE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = vowel)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )
```

## Mini-task 5: Change the Labels and Titles

-   Give your plot from Task 4 a new title!
-   Try and change the labels of the x and y axes too
    -   Previous plot code
        +`labs(title="Plot Title",caption="Plot Caption",x="column_a",y="column_b", color="column_c",fill="column_d")`

## Themes

-   You can change the background of the chart
    -   Grey squares is the default
-   Other options include:
    -   `+theme_bw()`, `+theme_light()`, `+theme_dark()`,
        `+theme_classic()`

## Chaning our theme

```{r,eval=FALSE, echo=TRUE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = vowel)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )+
  theme_bw()
```

## Themes

```{r,eval=TRUE, echo=FALSE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = vowel)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )+
  theme_bw()
```

## Mini-task 6: Change the Theme!

-   Change the theme for the plot from Task 5
    -   `+theme_bw()`
    -   `+theme_light()`
    -   `+theme_dark()`
    -   `+theme_classic()`

## Legends: positioning

-   The default legend position is to the right of the plot
-   It can be changed using `+theme(legend.position="")`
    -   Options are "bottom", "top", "left", or "right"

## Move our legend

```{r,eval=FALSE, echo=TRUE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = vowel)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )+
  theme_bw()+
  theme(legend.position = "left")
  
```

## Legends: positioning

```{r,eval=TRUE, echo=FALSE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = vowel)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )+
  theme_bw()+
  theme(legend.position = "left")
  
```

## Altering the legend

-   You can also alter the legend name and labels using:
    -   `scale_color_discrete(name="",labels=c("",""))` and
    -   `scale_fill_discrete(name="",labels=c("",""))`

## Altering our legend

```{r,eval=FALSE, echo=TRUE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = participant_gender)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )+
  theme_bw()+
  theme(legend.position = "left")+
  scale_color_discrete(name="Gender", 
                       labels=c("Female","Male"))
```

## Legends

```{r,eval=TRUE, echo=FALSE}
ggplot(data = qb_vowels, aes(x = F2_50, y = F1_50, colour = participant_gender)) +
  geom_point() +
  labs(
    title = "NZE Monophthongs",
    caption = "Data from QuakeBox Corpus",
    x = "Vowel frontness",
    y = "Vowel Height",
    color = "Vowel"
  )+
  theme_bw()+
  theme(legend.position = "left")+
  scale_color_discrete(name="Gender", 
                       labels=c("Female","Male"))
  
```

## Remove the legend

-   You can remove one of the legends if you have more than one with
    `guides(fill="none")` or `guides(color="none")`
    -   Need to specify the mapping you want to remove

## Removing the legend

```{r,eval=FALSE, echo=TRUE}
qb_vowels %>% 
  filter(vowel %in% c("DRESS","KIT")) %>% 
ggplot(aes(x=vowel, y=F1_50,fill=participant_age_category,color=participant_gender))+
  geom_boxplot()+
  labs(title="Height of NZE DRESS and KIT",
      caption="Data sfrom QuakeBox", 
      x="Vowel",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_discrete(name="Gender", 
                       labels=c("Women","Men"))+
  guides(fill = "none") +
  theme_bw()
```

## Remove the legend

```{r,eval=TRUE, echo=FALSE}
qb_vowels %>% 
  filter(vowel %in% c("DRESS","KIT")) %>% 
ggplot(aes(x=vowel, y=F1_50,fill=participant_age_category,color=participant_gender))+
  geom_boxplot()+
  labs(title="Height of NZE DRESS and KIT",
      caption="Data from QuakeBox", 
      x="Vowel",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_discrete(name="Gender", 
                       labels=c("Women","Men"))+
  guides(fill = "none") +
  theme_bw()
  
```

## Remove all legends

-   If you want to remove *all* legends you can use
    `+ theme(legend.position="none")`

```{r,eval=FALSE, echo=TRUE}
qb_vowels %>% 
  filter(vowel %in% c("DRESS","KIT")) %>% 
ggplot(aes(x=vowel, y=F1_50,fill=participant_age_category,color=participant_gender))+
  geom_boxplot()+
  labs(title="Height of NZE DRESS and KIT",
      caption="Data from QuakeBox", 
      x="Vowel",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_discrete(name="Gender", 
                       labels=c("Women","Men"))+
  guides(fill = "none") +
  theme_bw()+
  theme(legend.position="none")
  
```

## Remove all legends

```{r,eval=TRUE, echo=FALSE}
qb_vowels %>% 
  filter(vowel %in% c("DRESS","KIT")) %>% 
ggplot(aes(x=vowel, y=F1_50,fill=participant_age_category,color=participant_gender))+
  geom_boxplot()+
  labs(title="Height of NZE DRESS and KIT",
      caption="Data from QuakeBox", 
      x="Vowel",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_discrete(name="Gender", 
                       labels=c("Women","Men"))+
  guides(fill = "none") +
  theme_bw()+
  theme(legend.position="none")
```

## Mini-task 7: Change the Legend

-   Change the labels for your legend
-   Try and remove one of them if you have two

## Faceting

-   You can break down your graph further by categorical variables with
    `facet_wrap()`, which automatically wraps graphs in a rectangle
    layout
-   `facet_wrap(~)` can take one or two arguments
    -   for one argument, it goes to the right of the \~
        `facet_wrap(~column_e)`
    -   for two arguments, one goes on either side of the \~
        `facet_wrap(column_e~column_f)`

## Faceting using facet_wrap()

```{r,eval=FALSE, echo=TRUE}
ggplot(data=qb_vowels, aes(x=F1_50,y=F2_50,color=vowel))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from Sydney Speaks",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_wrap(participant_gender~participant_nz_ethnic)+
  scale_y_reverse()+
  scale_x_reverse()
```

## Faceting using facet_wrap()

```{r,eval=TRUE, echo=FALSE}
ggplot(data=qb_vowels, aes(x=F1_50,y=F2_50,color=vowel))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from Sydney Speaks",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_wrap(participant_gender~participant_nz_ethnic)+
  scale_y_reverse()+
  scale_x_reverse()
```

## Faceting with facet_grid()

-   `facet_grid()` can facet into columns or rows, or both
    -   `facet_grid(cols=vars(column_e))` will facet into columns based
        on this variable
    -   `facet_grid(rows=vars(column_f))` will facet into rows based on
        this variable
    -   `facet_grid(cols=vars(column_e), rows=vars(column_f))` will
        facet into rows and columns based on the two variables

## Faceting with facet_grid()

```{r,eval=FALSE, echo=TRUE}
ggplot(data=qb_vowels, aes(x=F1_50,y=F2_50,color=vowel))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from Sydney Speaks",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_grid(cols=vars(participant_gender),rows=vars(participant_nz_ethnic))+
  scale_y_reverse()+
  scale_x_reverse()
```

## Faceting with facet_grid()

```{r,eval=TRUE, echo=FALSE}
ggplot(data=qb_vowels, aes(x=F1_50,y=F2_50,color=vowel))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from Sydney Speaks",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_grid(cols=vars(participant_gender),rows=vars(participant_nz_ethnic))+
  scale_y_reverse()+
  scale_x_reverse()
```

## Faceting tip

-   By default, x and y axes scales are the same across facets
-   You can adjust them to be 'free' with the `scales` argument
    -   `scales = "free"`: free x and y axes
    -   `scales ="free_x"`: free x axes only
    -   `scales ="free_y"`: free y axes only

## Facet: default

```{r eval=FALSE}
qb_vowels %>% 
  filter(vowel %in% c("START","DRESS")) %>% 
ggplot(data=, aes(x=F1_50,y=F2_50,color=participant_gender))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from QuakeBox",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_wrap(~vowel)+
  scale_y_reverse()+
  scale_x_reverse()
```

## Facet: default

```{r echo: false}
qb_vowels %>% 
  filter(vowel %in% c("START","DRESS")) %>% 
ggplot(data=, aes(x=F1_50,y=F2_50,color=participant_gender))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from QuakeBox",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_wrap(~vowel)+
  scale_y_reverse()+
  scale_x_reverse()
```

## Facet: free scales

```{r eval=F}
qb_vowels %>% 
  filter(vowel %in% c("START","DRESS")) %>% 
ggplot(data=, aes(x=F1_50,y=F2_50,color=participant_gender))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from QuakeBox",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_wrap(~vowel, scales = "free")+
  scale_y_reverse()+
  scale_x_reverse()
```

## Facet: free scales

```{r echo: FALSE}
qb_vowels %>% 
  filter(vowel %in% c("START","DRESS")) %>% 
ggplot(data=, aes(x=F1_50,y=F2_50,color=participant_gender))+ 
  geom_point()+ 
  labs(title="NZE Monophthongs", 
       caption="Data from QuakeBox",
       x="Vowel Backness",y="Vowel Height",
       color="Vowel")+ 
  theme_bw()+
  facet_wrap(~vowel, scales = "free")+
  scale_y_reverse()+
  scale_x_reverse()
```

## Mini-Task 8: Facet it!

-   Try faceting your graph from task 7 by one or two categorical
    variables
-   Try both facet_wrap and facet_grid - which do you prefer?
    -   `facet_wrap(~column_e)`
    -   `facet_wrap(column_e~column_f)`
    -   `facet_grid(cols=vars(column_e), rows=vars(column_f))`

## Changing the colours

-   When you map variables onto colour, R will automatically select
    colours for you
    -   But often it's not cute or doesn't match the colour scheme of
        your presentation
-   ggplot lets us change the colours. You can:
    -   Manually select each colour within the plot
    -   Choose a pre-existing palette

## Manually selecting colours

-   Get a list of the available colour names with `colors()`
-   Or you can use the Hex code for the colour
-   To change the colours in a plot manually we add
    `scale_color_manual()`
    -   Inside the () we add `values=c()` with the colour names or hex
        codes in ""

## Manually selecting colours

```{r, eval=FALSE}
ggplot(data=qb_vowels, aes(x=F2_50,y=F1_50, color=participant_gender))+
  geom_point()+
  scale_color_manual(values=c("blue2","chartreuse2"))+
  scale_y_reverse()+
  scale_x_reverse()
```

## Manually selecting colours

```{r,echo=FALSE}
ggplot(data=qb_vowels, aes(x=F2_50,y=F1_50, color=participant_gender))+
  geom_point()+
  scale_color_manual(values=c("blue2","chartreuse2"))+
  scale_y_reverse()+
  scale_x_reverse()
```

## Manually selecting colours

-   Where you have both fill and colour, the same applies\
-   To change the fill colour manually add:
    -   `+scale_fill_manual()`
-   To change the line colour manually add:
    -   `+scale_color_manual()`

## Manually selecting colours

```{r,eval=FALSE}
qb_vowels %>% 
  filter(vowel %in% c("KIT","DRESS")) %>% 
ggplot(data=, aes(x=vowel, y=F1_50,fill=participant_nz_ethnic,color=participant_gender))+
  geom_boxplot()+
  labs(title="Realisation of vowel height by gender and ethnicity",
      caption="Data from QuakeBox", 
      x="Generation",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_manual(values=c("#C70039","#581845"),name="Gender",labels = c("Women","Men"))+
  scale_fill_manual(values=c("#DAF7A6","#FFC300","#FF5733","#C70039"),
  name="Ethnicity", 
  labels = c("NZ European","NZ Mixed","Other")
  )
```

## Manually selecting colours

```{r, echo=FALSE}
qb_vowels %>% 
  filter(vowel %in% c("KIT","DRESS")) %>% 
ggplot(data=, aes(x=vowel, y=F1_50,fill=participant_nz_ethnic,color=participant_gender))+
  geom_boxplot()+
  labs(title="Realisation of vowel height by gender and ethnicity",
      caption="Data from QuakeBox", 
      x="Generation",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_manual(values=c("#C70039","#581845"),name="Gender",labels = c("Women","Men"))+
  scale_fill_manual(values=c("#DAF7A6","#FFC300","#FF5733","#C70039"),
  name="Ethnicity", 
  labels = c("NZ European","NZ Mixed","Other")
  )
```

## Pre-existing palettes
-   You can assign multiple colours to an object to create a palette
    - e.g.,     `my_palette <- c("#DAF7A6","#FFC300","#FF5733","#C70039")`
- This object can be used as the `values` value
    -   e.g., `scale_color_manual(values=my_palette)`

## Pre-existing palettes

-   You don't need to go through choosing colours manually if you don't
    want to\
-   There are a lot of pre-existing palettes that you can add to your
    graph\
-   Different packages have different palettes, I use RColorBrewer\
-   We replace the 'manual' in the previous code with `brewer`:
    -   `scale_color_brewer(palette="")`\
    -   `scale_fill_brewer(palette="")`\
-   You can find palette names in the link in the script

## Pre-existing palettes

```{r, eval=FALSE}
qb_vowels %>% 
  filter(vowel %in% c("KIT","DRESS")) %>% 
ggplot(data=, aes(x=vowel, y=F1_50,fill=participant_nz_ethnic,color=participant_gender))+
  geom_boxplot()+
  labs(title="Realisation of vowel height by gender and ethnicity",
      caption="Data from QuakeBox", 
      x="Generation",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_brewer(palette="RdGy",name="Gender",labels = c("Women","Men"))+
  scale_fill_brewer(palette="Spectral",
  name="Ethnicity", 
  labels = c("NZ European","NZ Mixed","Other")
  )
```

## Pre-existing palettes

```{r, echo=FALSE}
qb_vowels %>% 
  filter(vowel %in% c("KIT","DRESS")) %>% 
ggplot(data=, aes(x=vowel, y=F1_50,fill=participant_nz_ethnic,color=participant_gender))+
  geom_boxplot()+
  labs(title="Realisation of vowel height by gender and ethnicity",
      caption="Data from QuakeBox", 
      x="Generation",y="Vowel Height")+
  theme(legend.position = "left")+
  scale_color_brewer(palette="RdGy",name="Gender",labels = c("Women","Men"))+
  scale_fill_brewer(palette="Spectral",
  name="Ethnicity", 
  labels = c("NZ European","NZ Mixed","Other")
  )

```

## Mini-task 9: Make it cute

-   Manually change the colours in a chart by changing the colours
    manually, or using a pre-existing palette \_ You can copy and paste
    a graph from above!

## Saving your figure with ggsave

-   It is possible to export figures from R Studio manually
-   `ggsave` gives more control over output
    -   Ensure higher quality
-   Can specify object, or it will save the most recently created figure

## Save your figure with ggsave

```{r eval=FALSE}
ggsave(
  filename = "file_path.png",
  dpi = 300,
  width = 500,
  height = 500,
  units = "px"
)
```

## Additional options

-   `Cowplot`
    -   Combine multiple figures together
-   `lemon`
-   `ggpatterns`
    -   Useful if no colour allowed
-   `ggalt`
    -   geom_encircle() function
-   `ggeffects`
    -   Plot model estimations as ggplot figures

## Ggplot competition: Bring it all together

-   Construct the worst plot you possibly can, using everything I've
    gone over today
-   Export it as a png
-   Email it to me at elena.sheard\@cantebury.ac.nz
-   Informal poll to decide the victor

## Troubleshooting

-   Remember to have colour and palette names/hex codes in " "
-   "" and () have to be closed
-   R is case-sensitive and space-sensitive
-   there must be a `+` connecting lines
-   Hex codes must have the \# before the digits

## Useful Resources

[R Graphics Cookbook](https://r-graphics.org/)\
[R Graph Gallery](https://www.r-graph-gallery.com/ggplot2-package.html)\
[ggplot2: Elegant Graphics for Data
Analysis](https://ggplot2-book.org/introduction.html)

### Color and palette generators

[Coolers](https://coolors.co/984447-faa6ff-64a6bd-ffba49-1e1e24)
[Colormind](http://colormind.io/) [Muzli](https://colors.muz.li/)
[Venngage](https://venngage.com/tools/accessible-color-palette-generator)
